---
title: "Reproducible Research: Peer Assessment 1"
author: "SR"
date: "8/29/2020"
output: html_document
keep_md: true
---

## Loading and preprocessing the data

Code to load data and process/transform the data into a format suitable for analysis

```{r echo = TRUE}
rawdata <- read.csv("activity.csv", header = TRUE)
totsteps <- aggregate(steps~date, data = rawdata, sum, na.rm = TRUE)
head(totsteps)
```

## What is mean total number of steps taken per day?

Histogram of the total number of steps taken each day

```{r echo = TRUE}
hist(totsteps$steps, col = "red", main = "Total number of steps per day", xlab = "Steps", breaks = 10)
```

The mean and median total number of steps taken per day

```{r echo = TRUE}
mean(totsteps$steps)
median(totsteps$steps)
```

## What is the average daily activity pattern?

Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r echo = TRUE}
avgsteps <- aggregate(steps~interval,data = rawdata, mean, na.rm = TRUE)
plot(steps~interval, data = avgsteps, type = "l", main = "Time series plot", xlab = "5-minute interval", ylab = "Average steps")
```

The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps

```{r echo = TRUE}
avgsteps[which.max(avgsteps$steps),]$interval
```

## Imputing missing values

The total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r echo = TRUE}
sum(is.na(rawdata$steps))
```

Code for filling in all of the missing values in the dataset using the mean and a new dataset that is equal to the original dataset but with the missing data filled in

```{r echo = TRUE}
library(magrittr)
library(dplyr)

replacena <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
newdata <- rawdata%>% group_by(interval) %>% mutate(steps= replacena(steps))
newtotsteps <- aggregate(steps~date, data = newdata, sum, na.rm = TRUE)
head(newtotsteps)
```

Histogram of the total number of steps taken each day using new data set with missing values replaced by mean.

```{r echo = TRUE}
hist(newtotsteps$steps, col = "red", main = "Total number of steps per day", xlab = "Steps", breaks = 10)
```

The mean and median total number of steps taken per day for new data

```{r echo = TRUE}
mean(newtotsteps$steps)
median(newtotsteps$steps)
```

The mean is the same as before but the median is slightly different from the previous value and the same as the mean


## Are there differences in activity patterns between weekdays and weekends?

Code for a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day

```{r echo = TRUE}
weeklydata <- newdata
weeklydata$date <- as.Date(strptime(weeklydata$date, format="%Y-%m-%d")) 
weeklydata$day <-  factor(ifelse(as.POSIXlt(weeklydata$date)$wday %in% c(0,6), 'weekend', 'weekday'))
```

Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r echo = TRUE}
avgweeklysteps <- aggregate(steps ~ interval + day, data = weeklydata, mean)
library(lattice)
xyplot(steps~interval|factor(day), data = avgweeklysteps, aspect = 1/2, type = "l", xlab = "5-minute interval", ylab = "Average steps", main = "Time series plot")
```

